---
title: Brazil data
author: Ricardo Mayer
date: '2018-01-16'
categories:
  - data
tags:
  - data
  - Brazil
slug: brazil-data
---

```{r setup, message=FALSE, echo=FALSE}
options(knitr.table.format = "html", knitr.kable.NA = '') 
library(kableExtra)
library(tidyr)
library(dplyr)
library(timetk)
library(readxl)
library(lubridate)
library(purrr)
library(forecast)
library(ggplot2)
library(tibbletime)
```


## Data sources and definition

## Data storage 

## Data exploration

### Data loading

load
```{r data_load}
country_q <- read_excel("../../Brasil.xlsx", sheet = "quarterly")
country_m <- read_excel("../../Brasil.xlsx", sheet = "monthly")
```



clean up

```{r cleanup,  message=FALSE, warning=FALSE, echo=FALSE}
# data_q <- country_q %>% select(-c(year, quarter, trim, hlookup)) 

data_q <- country_q %>% select(-c(year, quarter, trim, hlookup)) %>%
  filter(!is.na(rgdp))

data_m <- country_m %>% select(-c(year, month,  hlookup)) 

data_q$date <- ymd(data_q$date)
data_m$date <- ymd(data_m$date)
```

Three versions of the data set for different purposes:
- multicolumn ts (mts) object 
- multicolumn xts object
- a tibble with list column, to apply functions for each series and store results as another list column (think estimation model, results, etc)

```{r reshape_and_nesting, message=FALSE, warning=FALSE, echo=FALSE}
start_year_q <- min(year(data_q$date))
start_year_m <- min(year(data_m$date))


data_q_mts <- tk_ts(data_q, start = start_year_q, frequency = 4)
data_q_xts <- tk_xts(data_q, date_var = date)
data_q_tbt <- as_tbl_time(data_q, index = date)
data_q_long <- data_q %>% gather(variable, value, -date)

data_q_nested <- data_q_long %>% group_by(variable) %>% 
  nest(.key = "data.tbl")
data_q_n_ts <- data_q_nested %>% 
  mutate(data.ts = map(.x   =  data.tbl,
                       .f   =  tk_ts,
                       start = start_year_q,
                       freq = 4
                       )
         )

data_m_mts <- tk_ts(data_m, start = start_year_m, frequency = 12)
data_m_xts <- tk_xts(data_m, date_var = date)
data_m_tbt <- as_tbl_time(data_m, index = date)
data_m_long <- data_m %>% gather(variable, value, -date)

data_m_nested <- data_m_long %>% group_by(variable) %>% 
  nest(.key = "data.tbl")
data_m_n_ts <- data_m_nested %>% 
  mutate(data.ts = map(.x   =  data.tbl,
                       .f   =  tk_ts,
                       start = start_year_m,
                       freq = 12
                       )
         )

```

Some plots

Quaterly variables
```{r plotq, message=FALSE, warning=FALSE}
plotbase_q <- ggplot(data = data_q_long,
                     aes(x = date, y = value, colour = variable))  

plot_q <- plotbase_q + geom_line() + 
  facet_wrap(~variable, scales = "free_y", ncol = 3) + 
  theme(legend.position="none")

plot_q
```


Monthly variables
```{r plotm, message=FALSE, warning=FALSE}
plotbase_m <- ggplot(data = data_m_long,
                     aes(x = date, y = value, colour = variable))  

plot_m <- plotbase_m + geom_line() + 
  facet_wrap(~variable, scales = "free_y", ncol = 3) +
  theme(legend.position="none") 

plot_m
```


## Summary of properties

### Dates ranges


Ranges for monthly variables

```{r date_ranges_m, echo=FALSE}

data_m_long <- data_m_long %>% 
  mutate(date_not_nas = date) 

data_m_long[ is.na(data_m_long$value)   ,"date_not_nas"] <- NA

minmaxdates_m <- data_m_long %>% group_by(variable) %>% 
  filter(!is.na(date_not_nas)) %>% 
  summarise(min_date = min(date_not_nas),
            max_date = max(date_not_nas)
            ) %>% 
  arrange(desc(max_date), min_date)

monthly_absolute_min <- min(minmaxdates_m$min_date)
monthly_absolute_max <- max(minmaxdates_m$max_date)
```


```{r date_ranges_q, echo=FALSE}

data_q_long <- data_q_long %>% 
  mutate(date_not_nas = date) 

data_q_long[ is.na(data_q_long$value)   ,"date_not_nas"] <- NA

minmaxdates_q <- data_q_long %>% group_by(variable) %>% 
  filter(!is.na(date_not_nas)) %>% 
  summarise(min_date = min(date_not_nas),
            max_date = max(date_not_nas)
            ) %>% 
  arrange(desc(max_date), min_date)

quarterly_absolute_min <- min(minmaxdates_q$min_date)
quarterly_absolute_max <- max(minmaxdates_q$max_date)


latest_rgdp <- minmaxdates_q %>% filter(variable == "rgdp")
latest_rgdp <- latest_rgdp[["max_date"]]

```



Ranges for montlhy and quarterly variables

- Latest real GDP date:  `r latest_rgdp` 
- Latest monthly date: `r monthly_absolute_max`  
- Latest quarterly date: `r quarterly_absolute_max`  
- Earliest monthly date: `r monthly_absolute_min`  
- Earliest quarterly date: `r quarterly_absolute_min`  

Monthly variables *strictly* more up-to-date than GDP:
```{r monthly_leads, echo=FALSE}
knitr::kable(dplyr::filter(minmaxdates_m, max_date > latest_rgdp),
             format = "html") %>% 
  kable_styling(bootstrap_options = "striped")
```


 
All monthly variables:
```{r table_dates_m, echo=FALSE}
knitr::kable(minmaxdates_m, format = "html") %>% 
  kable_styling(bootstrap_options = "striped")

```


All quarterly variables
```{r table_dates_q, echo=FALSE}
knitr::kable(minmaxdates_q, format = "html") %>% 
  kable_styling(bootstrap_options = "striped")

```

plots with dates

```{r plot_dates_m, echo=FALSE}

p_date_base <- ggplot(data = data_m_long, 
                      aes(date_not_nas, variable)) 

p_date_m <- p_date_base + geom_line(aes(group = variable), na.rm = TRUE, col = "blue")
p_date_m

```



```{r plot_dates_q, echo=FALSE}

p_date_base_q <- ggplot(data = data_q_long, 
                      aes(date_not_nas, variable)) 

p_date_q <- p_date_base_q + geom_line(aes(x = date_not_nas, group = variable), na.rm = TRUE)
p_date_q

```
